\center Towards Globally Color-Dual All-Loop Integrands

### 2023-08-17 notes/2loopNLSM.tex
Added to Review section

\paragraph{YZ-Theory} Thus far, the two color-dual Lagrangians we have discussed make explicit reference to the spacetime dimension ($D=4$ for SDYM, and $D=2$ for ZM theory). This makes them poorly suited for constructing dimensionally regulated integrals in loop-level amplitudes. Here we briefly review YZ-theory, which is a honest $D$-dimensional theory first proposed by Cheung and Hsien \cite{Cheung2016prv} for the purpose of describing nonlinear sigma model (NLSM) scattering. One can understand YZ-theory as a particular dimensional reduction of Yang-Mills in $D=2d+1$ dimensions, down to $d$ dimensions.

Starting with the Yang-Mills Lagrangian, one can redefine the gauge fields, $A_M$, in terms of $X$, $Y$ and $Z$ fields:
\begin{align}
X_M &= (X_\mu,0,-iX_\mu) 
\\
Y_M &= (0,Y,0) 
\\
Z_M &= (Z_\mu,0,iZ_\mu) 
\end{align}
where $X$ and $Z$ are $d$-vectors and $Y$ is a scalar field. Considering the conjugate nature of the X-Z propagator, in this work we will make the replacement $X\rightarrow \bar{Z}$. Plugging this redefinition of the gauge fields in, upto cubic order, we thus obtain the following Lagrangian:
\begin{eBox}
\be
\mathcal{L}^{\text{YZ}} =\frac{1}{2} (\partial Y)^2 + (\partial Z)(\partial \bar{Z}) - g f^{abc} \left( \bar{Z}_{\mu\nu}Z^{\mu} Z^\nu + [Y,\partial_\mu Y] Z^\mu \right)
\ee
\end{eBox}
The Feynman rules for this Lagrangian are simply:
\be
\Acubic{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{} = i \varepsilon_3 (p_1-p_2)\qquad \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{} =i(\varepsilon_1 p_{2})(\varepsilon_2\bar{\varepsilon}_3) - i(\varepsilon_2 p_{1})(\varepsilon_1\bar{\varepsilon}_3)
\ee
Now we review of the construction of NLSM numerators at tree-level. In this construction, we can define generators of the kinematic algebra can be expressed as follows
\be\label{eq:FeynmanRuleYYZ}
T^a_{ij}= i \varepsilon_a(p_i-p_j)
\ee
where momentum conservation requires
\be
p_a + p_i + p_j =0
\ee
The kinematic half-ladder diagrams then takes on the following concise definition:
\be
n^{\text{NLSM}}_{(i|a_1a_2...a_n|j)} = (T^{a_1}T^{a_2}\cdots T^{a_n})_{ij}
\ee
Since there are on pole cancelling factors of $s_{ij} = (p_i+p_j)^2$, this definition of the chiral current algebra is manifestly cubic. Thus, the kinematic structure constants defined in terms of these generators are invariant under generalized gauge freedom. They can be defined implicitly below:
\be
[T^a,T^b]_{ij}= F^{a}_{\,b|c}T^c_{ij}
\ee
Given this definition, the Feynman rule associated with kinematic structure constant is simply:
\be\label{eq:FeynmanRuleXZZ}
i F^{a}_{\,b|c} = (\varepsilon_b p_{ab})(\varepsilon_a\bar{\varepsilon}_c) - (\varepsilon_a p_{ab})(\varepsilon_b\bar{\varepsilon}_c) 
\ee
where $p_{ab}=p_a+p_b$ and $\epsilon$ and $\bar{\epsilon}$ are the polarizations of the $Z$-vectors particle and its conjugate field, respectively. The on-shell state sum for the polarization vectors is simply:
\be
\sum_{\text{states}} \varepsilon^{\,\mu}_{(p)}\bar{\varepsilon}^{\,\nu}_{(-p)} = \eta^{\mu\nu}
\ee
Notice that the vector state sum is gauge fixed since the $Y\,Z$ model explicitly chooses Lorenz gauge for the $Z$ particles, $\partial_\mu Z^\mu=0$. To recover NLSM amplitudes from this these kinematic structure constants at tree-level, we simply plug in the following on-shell polarizations for the $Z$ and $\bar{Z}$ particles:
\be\label{eq:onShellZStates}
\varepsilon^\mu_{(p)} = p^\mu \qquad \bar{\varepsilon}^\mu_{(p)} = \frac{q^\mu}{pq}
\ee
where $q^2=0$ is some null reference momentum. With this, we can define tree-level pion scattering in two equivalent ways:
\be
A^{\text{NLSM}} = A(...,Y,...,Y,...) = A(...,\bar{Z},...) 
\ee
where the ellipses denote additional on-shell $Z$-particles. Subject to a particular gauge choice, the kinematic numerators in latter definition for pion scattering is equivalent to those $J$-theory, first written down by Cheung and one of the authors \cite{Cheung:2021zvb}. In this paradigm, the on-shell $\bar{Z}$ state is equivalent to the root leg appearing in the color-dual $J$-theory numerators.

@@

@@

\paragraph{ZM-theory}
While J-theory manifests color-kinematics duality, it only does so at tree level because J-theory breaks manifest Bose symmetry.
Manifest color-kinematics duality can be achieved at all loop orders by restricting J-theory to two spacetime dimensions where the chiral current can be dualized to a scalar $J^\mu = \epsilon^{\mu\nu} \partial_\nu \varphi$.
This trivially enforces Lorenz gauge while the vanishing field strength condition...

The Lagrangian for Zakharov-Mikhailov (ZM) theory is a simple $SU(N)$ scalar theory defined in two spacetime dimensions:
\begin{eBox}
\be
\mathcal{L}^{\text{ZM}} = \frac{1}{2}(\partial \varphi)^2 + g f^{abc} \varphi^a (\partial^\mu \varphi^b)( \tilde{\partial}_\mu \varphi^c) =  \text{tr}\left[\frac{1}{2}(\partial \varphi)^2-i g \varphi [\partial_t \varphi, \partial_z \varphi ]\right]
\ee
\end{eBox}
where $\tilde{\partial}_\mu = \epsilon_{\mu\nu}\partial^\nu $ and $\epsilon_{\mu\nu}$ is the two-dimensional Levi-Civita symbol. Expanded out into spacetime components, the momentum space Feynman rule for this theory is simply:
\be
\langle ab\rangle \equiv p_a^0 p_b^1 - p_a^1 p_b^0 = -\langle ba \rangle 
\ee
 From the interaction part of the Lagrangian, we can see that the three-point vertex is (1) functionally symmetric and (2) the kinematics are completely anti-symmetric: 
\be
\cubic{hgrey0}{}{nhpBlue}{}{nhpBlue}{}{nhpBlue} = \langle 12\rangle = \langle 23\rangle = \langle 31\rangle = -\langle 21\rangle
\ee
This important property is due to momentum conservation alone and does not need any additional kinematic constraints to satisfy the duality. In addition to anti-symmetry, the cubic graphs will satisfy jacobi off-shell on any internal edge due to the delta function identity of 2D Levi-Civita symbols (TODO:  show the identity, say that the following is Schouten, cite Schouten?, comment on addition of masses):
\be
\langle ab \rangle \langle cd\rangle +\langle ac \rangle \langle db\rangle +\langle ad \rangle \langle bc\rangle = \delta^{[a}_c\delta^{b]}_d + \delta^{[a}_d\delta^{c]}_b+ \delta^{[a}_b\delta^{d]}_c=0
\ee
Theories with this property are the creme de la creme of color-dual gauge theories as they trivially satisfy color-kinematic algebraic relations off-shell. This allows one to trivially promote the duality from tree-level to all-loop order by dressing each cubic vertex with its own kinematic structure constant.

@@

@@

\item I believe the j-theory Lagrangian does not lead to the correct EOM as currently written.

>>>>

\paragraph{YZ-Theory}
SDYM makes explicit reference to the spacetime dimension and Zakharov-Mikhailov theory, to be discussed in \sect{sec:ZMTheory}, does as well.
This makes both theories poorly suited for constructing dimensionally-regulated loop-level amplitudes.
On the other hand, YZ-theory is an honest $D$-dimensional theory that captures the classical physics of the nonlinear sigma model (NLSM) while manifest the duality between color and kinematics \cite{Cheung2016prv}.

One can understand YZ-theory as a particular dimensional reduction of Yang-Mills in $D=2d+1$ dimensions, down to $d$ dimensions. 
Starting with the Yang-Mills Lagrangian, one can redefine the gauge fields, $A_M$, in terms of $X$, $Y$ and $Z$ fields:
\begin{align}
X_M &= (X_\mu,0,-iX_\mu) 
\\
Y_M &= (0,Y,0) 
\\
Z_M &= (Z_\mu,0,iZ_\mu) 
\end{align}
where $X$ and $Z$ are $d$-vectors and $Y$ is a scalar field. Considering the conjugate nature of the X-Z propagator, in this work we will make the replacement $X\rightarrow \bar{Z}$. Plugging this redefinition of the gauge fields in, upto cubic order, we thus obtain the following Lagrangian:
\begin{eBox}
\be
\mathcal{L}^{\text{YZ}} =\frac{1}{2} (\partial Y)^2 + (\partial Z)(\partial \bar{Z}) - g f^{abc} \left( \bar{Z}_{\mu\nu}Z^{\mu} Z^\nu + [Y,\partial_\mu Y] Z^\mu \right)
\ee
\end{eBox}
The Feynman rules for this Lagrangian are simply:
\be
\Acubic{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{} = i \varepsilon_3 (p_1-p_2)\qquad \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{} =i(\varepsilon_1 p_{2})(\varepsilon_2\bar{\varepsilon}_3) - i(\varepsilon_2 p_{1})(\varepsilon_1\bar{\varepsilon}_3)
\ee
Now we review of the construction of NLSM numerators at tree-level. In this construction, we can define generators of the kinematic algebra can be expressed as follows
\be\label{eq:FeynmanRuleYYZ}
T^a_{ij}= i \varepsilon_a(p_i-p_j)
\ee
where momentum conservation requires
\be
p_a + p_i + p_j =0
\ee
The kinematic half-ladder diagrams then takes on the following concise definition:
\be
n^{\text{NLSM}}_{(i|a_1a_2...a_n|j)} = (T^{a_1}T^{a_2}\cdots T^{a_n})_{ij}
\ee
Since there are on pole cancelling factors of $s_{ij} = (p_i+p_j)^2$, this definition of the chiral current algebra is manifestly cubic. Thus, the kinematic structure constants defined in terms of these generators are invariant under generalized gauge freedom. They can be defined implicitly below:
\be
[T^a,T^b]_{ij}= F^{a}_{\,b|c}T^c_{ij}
\ee
Given this definition, the Feynman rule associated with kinematic structure constant is simply:
\be\label{eq:FeynmanRuleXZZ}
i F^{a}_{\,b|c} = (\varepsilon_b p_{ab})(\varepsilon_a\bar{\varepsilon}_c) - (\varepsilon_a p_{ab})(\varepsilon_b\bar{\varepsilon}_c) 
\ee
where $p_{ab}=p_a+p_b$ and $\epsilon$ and $\bar{\epsilon}$ are the polarizations of the $Z$-vectors particle and its conjugate field, respectively. The on-shell state sum for the polarization vectors is simply:
\be
\sum_{\text{states}} \varepsilon^{\,\mu}_{(p)}\bar{\varepsilon}^{\,\nu}_{(-p)} = \eta^{\mu\nu}
\ee
Notice that the vector state sum is gauge fixed since the $Y\,Z$ model explicitly chooses Lorenz gauge for the $Z$ particles, $\partial_\mu Z^\mu=0$. To recover NLSM amplitudes from this these kinematic structure constants at tree-level, we simply plug in the following on-shell polarizations for the $Z$ and $\bar{Z}$ particles:
\be\label{eq:onShellZStates}
\varepsilon^\mu_{(p)} = p^\mu \qquad \bar{\varepsilon}^\mu_{(p)} = \frac{q^\mu}{pq}
\ee
where $q^2=0$ is some null reference momentum. With this, we can define tree-level pion scattering in two equivalent ways:
\be
A^{\text{NLSM}} = A(...,Y,...,Y,...) = A(...,\bar{Z},...) 
\ee
where the ellipses denote additional on-shell $Z$-particles. Subject to a particular gauge choice, the kinematic numerators in latter definition for pion scattering is equivalent to those $J$-theory, first written down by Cheung and one of the authors \cite{Cheung:2021zvb}. In this paradigm, the on-shell $\bar{Z}$ state is equivalent to the root leg appearing in the color-dual $J$-theory numerators.

@@

\label{eq:JTheoryEOM}

@@

\draftnoteJM{Put comment about the fact that you can use funny propagators without affecting CK earlier.  Mention NANS as modification of J-theory.}

\paragraph{ZM-theory}
\label{sec:ZMTheory}
While J-theory manifests color-kinematics duality, it only does so at tree level because J-theory breaks manifest Bose symmetry.
Manifest color-kinematics duality can be achieved at all loop orders by restricting J-theory to two spacetime dimensions where the chiral current can be dualized to a scalar $J^\mu = \epsilon^{\mu\nu} \partial_\nu \varphi$, which trivially enforces Lorenz gauge.
After plugging $\varphi$ into \eqref{eq:JTheoryEOM} and rearranging, another copy of $\varphi$ (rather than some new field) can be integrated in to obtain the Zakharov-Mikhailov (ZM) Lagrangian
\begin{eBox}
\be
\mathcal{L}^{\text{ZM}} = \frac{1}{2}(\partial \varphi)^2 + g f^{abc} \varphi^a (\partial^\mu \varphi^b)( \tilde{\partial}_\mu \varphi^c) =  \text{tr}\left[\frac{1}{2}(\partial \varphi)^2-i g \varphi [\partial_t \varphi, \partial_z \varphi ]\right]
\ee
\end{eBox}
where $\tilde{\partial}_\mu = \epsilon_{\mu\nu}\partial^\nu $.
This procedure makes it clear that ZM and NLSM encode the same classical physics, that is, their equations of motion are dual to each other.
The color-stripped Feynman rule for the three-point vertex can be read off from the Lagrangian, giving
\be
\langle ab\rangle \equiv p_a^\mu \epsilon_{\mu\nu} p_b^\nu =p_a^0 p_b^1 - p_a^1 p_b^0 = -\langle ba \rangle .
\ee
The vertex is (1) functionally symmetric and (2) the kinematic portion of the vertex is completely anti-symmetric: 
\be
\cubic{hgrey0}{}{nhpBlue}{}{nhpBlue}{}{nhpBlue} = \langle 12\rangle = \langle 23\rangle = \langle 31\rangle = -\langle 21\rangle
\ee
This important property is due to momentum conservation alone and does not need any additional kinematic constraints to satisfy the duality.
To prove off-shell color kinematics duality for any multiplicity and loop order it is enough to observe that the sum of the off-shell $s$-, $t$-, and $u$-channel numerators sum to zero
\be
\label{eq:ZMJacobi}
\langle ab \rangle \langle cd\rangle +\langle ac \rangle \langle db\rangle +\langle ad \rangle \langle bc\rangle =0 % \delta^{[a}_c\delta^{b]}_d + \delta^{[a}_d\delta^{c]}_b+ \delta^{[a}_b\delta^{d]}_c
\ee
by virtue of the 2D Levi-Civita identity $\epsilon_{\mu\nu}\epsilon_{\rho\sigma} = \eta_{\mu\rho}\eta_{\nu\sigma}-\eta_{\mu\sigma}\eta_{\nu\rho}$.
Amusingly, \eqref{eq:ZMJacobi} is simply a manifestation of the Schouten identity.

Since ZM theory manifested color-kinematics duality without any reference to the on-shell condition, a mass term can be added to ZM theory without spoiling color-kinematics duality.
In fact, the kinetic term could be dramatically altered.
In general, if the on-shell conditions are not used, color-kinematics duality admits any kinetic term but the choice of propagator might spoil locality in the double-coppied theory.

@@

\paragraph{Generalities of known color-dual theories -- James (I'll come back to this later)}
\begin{itemize}
\item There's also CS but the amplitudes are all zero (and marginally 10D N=1 SYM)
\item All of the theories described above have diff as kinematic algebra
\item All of the theories choose a simple gauge like Lorenz or lightcone
\item All of the theories place some constraint on the field strength directly and get their equations of motion directly from that constraint, not the YM EOM.
\item (Almost all of the theories (aside from NLSM) have trivial amplitudes at least at tree level)
\item This means that almost all of the theories only have CK at the classical level, which can sometimes be fudged into 1-loop as we'll now discuss.
\end{itemize}

@@
